Fast Sorting on Modern Computers

### 2005-06-24 docs/tech_report/heap.tex
made the thing compile now
managed to compile the docs, without the 16 byte cache line

\section{Results}
All simulations presented here were performed on 32 byte cache lines. As a
result, 8-heaps were used for heapsort, since 8 4-byte \n{int}s fit the cache line.
LaMarca used 64 bit ints, and while this doesn't affect results, it means that
fewer keys fit into his cache lines. As a result, he used a 4-heap. To try and
emulate his results, extra tests were done using a 4-heap with a 32 byte
cache line, and using a 4-heap using a 16 byte cache line, so that the entire
cache line is used. \verb!UINT_MAX! was used as the sentinel.

@@

An additional consideration is that the base heapsort has more instructions than
it's out-of-place equivalent. This means that the 4-heap will have a better
instruction count, as well as better cache performance than the base heapsort.
The 8-heap should have the best cache performance of them all, due to fully
using each cache line, taking advantage of the spatial locality of the cache
line. However, the 4-heap should perform just as well with a 16 byte cache line.

@@

\plot{heap}{}{}{Cache heapsort (4-heap) was also tested with a 16 byte cache line}{Cache heapsort (4-heap) was also tested with a 16 byte cache line}{}{}

@@

\sixplot{base_heapsort-0.eps}{cache4_heapsort-0.eps}{cache_heapsort-0.eps}{cache_heapsort-1.eps}{}{}{base and cache heapsorts}

@@

\begin{figure}
\includegraphics{../counters/curve.eps}
\caption{Predictability of branches in cache heapsort, with a fanout of eight}
\label{cache heapsort curve}
\end{figure}

@@

Comparing the 4-heap with a 16 byte cache line to the 4-heap with a 32 byte
cache line is not as interesting as it would be to compare it to an 8-heap on a
16 byte cache line.

Also, it would be interesting to use performance counters to find out whether
the cache or the branch predictors are responsible for the base heapsort being
faster than the 4-heap, on the Pentium 4.

>>>>

\section{Results}
All simulations presented here were performed on 32 byte cache lines. As a
result, 8-heaps were used for heapsort, since 8 4-byte \n{int}s fit the cache line.
LaMarca used 64 bit ints, and while this doesn't affect results, it means that
fewer keys fit into his cache lines. Since modern cache lines are typically
longer than 4 bytes, we stick with 8-byte cache lines. \verb!UINT_MAX! was used
as the sentinel.

@@

An additional consideration is that the base heapsort has more instructions than
it's out-of-place equivalent. This means that the 4-heap will have a better
instruction count, as well as better cache performance than the base heapsort.
The 8-heap should have the best cache performance of them all, due to fully
using each cache line, taking advantage of the spatial locality of the cache
line.

@@

\plot{heap}{}{}{}{}{}{}

@@

\sixplot{base_heapsort_0.eps}{cache4_heapsort_0.eps}{cache_heapsort_0.eps}{cache_heapsort_1.eps}{}{}{base and cache heapsorts}

@@

\begin{figure}
\includegraphics{plots/curve.eps}
\caption{Predictability of branches in cache heapsort, with a fanout of eight}
\label{cache heapsort curve}
\end{figure}

@@
